\subsection[Quiz]{Week 3 Quiz}
\subsubsection*{1. Which of the following are true?}
\paragraph*{The degree of a finite extension $K / \F_p$ is divisible by $p$.}
This claim is \emph{false}: for example, take $\F_2[x]/(x^3 + x + 1)$ over $\F_2$. 
Since the polynomial has degree $3$ the extension has order $2^3 = 8$; but the \emph{degree} is $3$ which is not divisible by $2$.

\paragraph*{The number of elements in a field extension $K / \F_p$ is equal to $p^n$ for some $n$.}
This claim is \emph{true}.

\paragraph*{$\F_{p^2}$ can be embedded as a subfield of $\F_{p^3}$.}
This claim is \emph{false} by Thm. 8: We can perform an embedding $\F_{p^d} \embed \F_{p^n}$ only if $d | n$, and $2$ does \emph{not} divide $3$.

\paragraph*{$\F_{p^2}$ can be embedded as a subfield of $\F_{p^4}$.}
This claim is \emph{true} by Thm. 8: Since $2$ divides $4$, we can perform the embedding.

\paragraph*{Any two finite fields of the same order are isomorphic.}
This claim is \emph{true}.

\subsubsection*{2. Which of the following are true?}
\paragraph*{A finite field of order $p^n$ has a unique subfield of order $p^m$ for every $m$ dividing $n$.}
This is \emph{true} by Thm. 8: For each $m$ dividing $n$, the theorem implies that $\F_{p^m} \sbs \F_{p^n}$.

\paragraph*{An algebraic closure $\overline{\F_p}$ can contain many different subfields of order $p^n$ for some $n$.}
This claim is \emph{false} by Thm. 7: such a subfield will be unique.

\paragraph*{For every $n$, there is an irreducible polynomial $P \in \F_p[x]$ of degree $n$}
This claim is \emph{true} by Cor. 7 (to Thm. 10).

\paragraph*{The group of automorphisms of a field $F$ of order $p^n$ has order $n!$.}
This claim is \emph{false}: the group is cyclic so it has order $p^n$.

\subsubsection*{3. Which of the following are true?}
\paragraph*{It is possible for an irreducible polynomial in $\Q[x]$ to be inseparable (have multiple roots).}
This claim is \emph{false}, because $\Q$ has characteristic $0$ and therefore irreducibility implies separability.

\paragraph*{Every irreducible polynomial over a finite field $K$ is separable (has no multiple roots).}
This claim is \emph{true}: see example 10.

\paragraph*{The polynomial $x^p - t \in \F_p(t)[x]$ over the field of rational functions $\F_p(t)$ is separable.}
This claim is \emph{false}: The derivative $P^\prime = (d/dx)(x^p - t) = p x^{p-1} = 0$, so $\gcd(P, P^\prime) = P \neq 1$. Therefore the polynomial is not separable.

\paragraph*{The stem field $F_p(t^{1/p})$ for the polynomial $x^p - t$ has degree $p$ over $\F_p(t)$ and is a splitting field for $x^p - t$.}
This claim is \emph{true}: $\F_p(t^{1/p})$ is generated by $1, t^{1/p}, t^{2/p}, \dotsc, t^{1 - 1/p}, t$ over $\F_p$, so it has degree $p$.

Define $\alpha = t^{1/p}$.

\paragraph*{}

\subsubsection*{4. Which of the following are true?}
\paragraph*{}

\paragraph*{}

\paragraph*{}

\paragraph*{}

\paragraph*{}